---
title: MIBWR - Probability - [2]
author: Duc Nguyen
date: '2022-07-25'
slug: mibwr-probability-2
categories:
  - xác suất thống kê
tags:
  - tài liệu
subtitle: ''
summary: ''
authors: []
lastmod: '2022-07-25T15:30:34+07:00'
featured: no
output:
  blogdown::html_page:
    toc: yes
weight: ~
bibliography: D:/GITHUB/v5/zoterobib/xacsuat.bib
link-citations: yes
nocite: |
  @R-bookdown
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

## Tài liệu tham khảo
Nguồn: [@crainiceanuMethodsBiostatistics]

We start by reading a small data set. First, let us tell R where the file is:

```{r}
file_name = file.path("D:", "GITHUB", "v5", "R", "dataset", "bmi_age.txt") # Nhập vị trí file dữ liệu
data_bmi = read.table(file = file_name, 
                      header = TRUE, # File contains a header row 
                      stringsAsFactors = FALSE) # We do not want to convert strings/words to a categorical vari-able/factor
print(data_bmi)
```

Check this dataset

```{r}
dim(data_bmi)

colnames(data_bmi) # PID là mã ID của đối tượng nghiên cứu; BMI là chỉ số cân nặng-chiều cao, SEX là giới tính, AGE là số tuổi.

mean(data_bmi$BMI)

sd(data_bmi$BMI)
```

We use the dollar sign `$` to indicate that we are extracting a single variable from the data_bmi dataset. 

For simplicity, we will assign all the columns of the dataset to their own separate variables using the `attach` function. 

We do not recommend this approach in general, especially when multiple datasets are being used, since variables across records can become de-linked and independently sorted. 

We will do this only for this data set to simplify the code example.

```{r}
attach(data_bmi)
```

We can also show a basic plot (fancier plotting options will be shown later) of age vs. BMI:

```{r}
plot(AGE, BMI,type="p", pch=20, cex=2, col="blue") # Vẽ đồ thị tương quan giữa AGE và BMI

rect(xleft = 45, xright = 100, ybottom = 0, ytop = 26, col = scales::alpha("purple", 0.5)) # tô màu vùng dữ liệu

rect(xleft = 45, xright = 100, ybottom = 26, ytop = 100, col = scales::alpha("red", 0.5)) # tô màu vùng dữ liệu

rect(xleft = 0, xright = 45, ybottom = 0, ytop = 26, col = scales::alpha("blue", 0.5)) # tô màu vùng dữ liệu

points(AGE, BMI, type="p", pch=20, cex=2, col="blue") # Đưa các điểm dữ liệu vào đồ thị (đồ đậm lại với màu xanh dương)

text(x = AGE + 1, # Vị trí của con số hiện kế bên cái chấm cách 1 khoảng là 1 đơn vị trên trục hoành, nhìn cho đẹp
     y = BMI, labels = PID, col = "black") # Ghi chú trục tung và trục hoành

abline(h = 26, v = 45, col = "black") # Vẽ hai đường ngang và dọc ở vị trí 26 và 45
```

**Sử dụng đồ thị này để giải thích các thuật ngữ về phép tính trên tập hợp**

Here the call to abline is for drawing vertical and horizontal lines so that we can count the numbers in each respective box. 

We now make the connection between set theory and operations clearer in the context of the data. 

In particular, we emphasize how set operations translate into logic operators that can then be used for data extraction and operations. 

Consider the following subsets of subjects: $A$ subjects with $BMI < 26$ and $B$ subjects older than 45 $(AGE> 45)$.

Construct a logical indicator for which records fall into $A$.

```{r}
index_BMI_less_26 = BMI < 26 # which is represented by the points in the purple/blue regions and B
```

```{r}
index_AGE_greater_45 = AGE > 45 # which is represented by the points in the purple/red regions
```

Display the IDs for A and B. 

Here PID is the unique patient ID, though in many applications the PID can be more complex than just integer numbers.

```{r}
PID[index_BMI_less_26]
```

```{r}
PID[index_AGE_greater_45]
```

Let us calculate ${(A \cap B)^c}$, the complement of the intersection between A and B, which is shown by the non-purple regions.

These are subjects who do not/are (note the $c$ that stands for complement) (have a BMI less than 26) and (older than 45) and

```{r}
index_A_int_B_compl = !(index_BMI_less_26 & index_AGE_greater_45)
PID[index_A_int_B_compl]
```

The <mark>translation</mark> of ${(A \cap B)^c}$ into R is `!(index_BMI_less_26 & index_AGE_greater_45)`.

Note that `!` indicates is `not`, or `complement`, and `&` indicates `and`, or `intersection`.

So, the resulting IDs are everybody, except the subject with IDs 4 and 7. 

It would be instructive to conduct the same type of analysis for ${A^c} \cup {B^c}$, ${\left( {A \cup B} \right)^c}$, and ${A^c}$. Đại khái là cách thực hiện tìm các phần tử của những tập hợp này cũng tương tự như trên khi sử dụng `R` để viết hàm.   

While this dataset is relatively small for educational purposes, similar subsetting approaches are both essential and *impossible* to do by hand in large datasets.

## An intuitive introduction to the bootstrap

A major problem in practice is that, even if we run two identical experiments, data are never the same. 

For example, if two large epidemiological studies collect data on the effects of smoking on developing lung cancer, and have the exact
same sampling protocol and length, then the studies will select different people, with different demographic information and different times of transition from being healthy to having lung cancer. 

However, even though the two samples will be different, they will have some things in common. 

Those things are the target of estimation, the probability and time for conversion from being healthy to developing lung cancer, and the original target population (e.g., the US population). 

In practice we often have one study and we are interested in understanding what would happen if multiple studies were conducted. 

The reason for that is fundamental and it has to do with the generalizability of the experiment. 

Indeed, a study would collect data for a subsample of the population to make predictions about the rate and expected time of transition to lung cancer in the overall population.

**Bootstrap is a widely used statistical technique based on resampling the data.**

The bootstrap is designed to create many potential studies that share the same characteristics with the study that collected the data and may represent the **variability** of running these experiments *without actually running the experiments*.




