---
title: "Data cleaning and time series 1"
author: "Duc Nguyen"
date: '2022-07-09'
slug: data-cleaning-and-time-series-1
categories: kỹ năng r
tags: case study
subtitle: ''
summary: ''
authors: []
lastmod: '2022-07-09T15:16:53+07:00'
featured: no
bibliography: D:/GITHUB/v5/zoterobib/xacsuat.bib
link-citations: yes
nocite: |
  @R-bookdown
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
toc: yes
type: book
---

## 0. Load packages

```{r results='hide', message=FALSE, warning=FALSE}
library(here)
library(lubridate) # For working with date and time
library(dplyr)
library(mice) # For imputation
library(VIM) # For imputation
library(stringr) # Find and replace strings in dataframe
```

## 1. Code for reading in the dataset and/or processing the data

```{r}
file_csv <- here("R", "dataset", "activity.csv") 
activity <- read.csv(file_csv)
activity$date <- ymd(activity$date) # Transform 'date' variable
```

## 2. Histogram of the total number of steps taken each day

```{r}
total_step <- activity %>% group_by(date) %>% summarise(total = sum(steps, na.rm = TRUE))
hist(total_step$total, breaks = 8, xlab = "Steps", main = "Total number of steps taken each day")
```

## 3. Mean and median number of steps taken each day

```{r}
mean_step_happy <- activity %>% group_by(date) %>% summarise(mean_happy = mean(steps, na.rm = TRUE))
plot(mean_happy ~ date, mean_step_happy, xlab = "Time", ylab = "Steps", main = "Mean number of steps taken each day")
```

```{r}
median_step_happy <- activity %>% group_by(date) %>% summarise(median_happy = median(steps, na.rm = TRUE))
plot(median_happy ~ date, median_step_happy, xlab = "Time", ylab = "Steps", main = "Median number of steps taken each day")
```

## 4 & 5. Time series plot of the average number of steps taken and the 5-minute interval that, on average, contains the maximum number of steps

In this plot, the black line represents for the average number of steps taken in each day (left y-axis) while the blue line is for the maximum steps at the 5-minute interval in a day (right y-axis).

```{r}
par(mar = c(5, 4, 4, 4) + 0.3)

plot(mean_happy ~ date, data = mean_step_happy, type = "l", col = "black", 
     xlab = "Time", ylim = c(0, 100), ylab = "Steps", main = "Time series of average number of steps taken")

# Subset the 5-minute interval with maximum steps in a day

max_interval <- activity %>% group_by(date) %>% filter(steps == max(steps))

par(new = TRUE) # Add new plot

plot(interval ~ date, data = max_interval, type = "l", col = "blue", xlab = "", ylab = "", ylim = c(0,2500), axes = FALSE)

axis(side = 4, at = pretty(range(max_interval$interval)))  

mtext("5-minute interval", side = 4, line = 3) 

legend("topright", lty = 1, col = c("black", "blue"), legend = c("Steps", "5-minute interval"), cex = 0.7, bg = "transparent")
```

# 6. Code to describe and show a strategy for imputing missing data

Solution 1: Show NAs in dataset

```{r}
activity_ok <- summary(activity)
activity_ok <- as.data.frame(activity_ok)
number_na_value_step <- as.numeric(gsub('.*:', '', activity_ok[7,3]))
mean(is.na(activity$steps))
```

Solution 2: Show NAs in dataset using `mice` and `VIM` packages.

```{r}
md.pattern(activity)
aggr(activity, prop = T, numbers = T)
```

I make imputation for this dataset using method `pmm` as Predictive mean matching.

```{r}
activity_steps <- mice(activity, seed = 2468, printFlag = F)
activity_imputed <- complete(activity_steps, 1) # Select the first version of imputation dataset
```

# 7. Histogram of the total number of steps taken each day after missing values are imputed

The impact of imputing missing data on the estimates of the total daily number of steps is to make the histogram less skewness and to become as standard distribution chart.

```{r}
total_step_imputed <- activity_imputed %>% group_by(date) %>% summarise(total = sum(steps, na.rm = TRUE))

par(mfrow = c(1, 2), mar = c(5, 4, 5, 2))

hist(total_step$total, breaks = 8, xlab = "Steps", main = "Before imputation", cex.main = 0.8)

hist(total_step_imputed$total, breaks = 8, xlab = "Steps", main = "After imputation", cex.main = 0.8)

mtext("Total number of steps taken each day", side = 3, line = -2, outer = TRUE)
```

Calculate and report the mean and median total number of steps taken per day. 

Do these values differ from the estimates from the first part of the assignment?

For mean values, the imputated values (shown in blue dots) seem fitting to the overal trend of the original mean graph. While median values does not change its value, they stay in zeros because a number of zero step taken in each day rather than step-recorded values.

```{r}
mean_step_imputed <- activity_imputed %>% group_by(date) %>% 
        summarise(mean_imputed = mean(steps, na.rm = TRUE))

ok_mean <- cbind(mean_step_happy, mean_step_imputed)
yes_1 <- subset(ok_mean, mean_happy == "NaN")
yes_mean <- yes_1 %>% select(date, mean_imputed)

plot(mean_imputed ~ date, mean_step_imputed, xlab = "Time", ylab = "Steps", main = "Mean number of steps taken each day", ylim = c(0, 110))
points(mean_imputed ~ date, data = yes_mean, col = "blue", pch = 16)
legend("topright", pch = c(21, 16), col = c("black", "blue"), legend = c("Original", "Imputation"), cex = 0.8, bg = "transparent")

```

```{r}
median_step_imputed <- activity_imputed %>% group_by(date) %>% 
        summarise(median_imputed = median(steps, na.rm = TRUE))

ok_median <- cbind(median_step_happy, median_step_imputed)
ok_median[is.na(ok_median)] <- "missing value"
yes_2 <- subset(ok_median, median_happy == "missing value")
yes_median <- yes_2 %>% select(date, median_imputed)

plot(median_imputed ~ date, median_step_imputed, xlab = "Time", ylab = "Steps", main = "Median number of steps taken each day")
points(median_imputed ~ date, data = yes_median, col = "blue", pch = 16)
legend("topright", pch = c(21, 16), col = c("black", "blue"), legend = c("Original", "Imputation"), cex = 0.8, bg = "transparent")
```

What is the impact of imputing missing data on the estimates of the total daily number of steps?

It helps connect the missing dots in the original chart thereby smoothing the line.

```{r}

plot(mean_imputed ~ date, data = mean_step_imputed, type = "l", col = "red", 
     xlab = "Time", ylim = c(0, 100), ylab = "Steps", main = "Time series of average number of steps taken")

lines(mean_happy ~ date, data = mean_step_happy, type = "l", col = "black", 
     xlab = "Time", ylim = c(0, 90), ylab = "Steps", main = "Time series of average number of steps taken")

legend("topright", lty = 1, col = c("black", "red"), legend = c("Original", "Imputation"), cex = 1, bg = "transparent")
```

# 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
activity_week <- activity_imputed %>% mutate(status = weekdays(date))

activity_week$status <- str_replace_all(activity_week$status, "Monday", "weekday") 
activity_week$status <- str_replace_all(activity_week$status, "Tuesday", "weekday")
activity_week$status <- str_replace_all(activity_week$status, "Wednesday", "weekday")
activity_week$status <- str_replace_all(activity_week$status, "Thursday", "weekday")
activity_week$status <- str_replace_all(activity_week$status, "Friday", "weekday")
activity_week$status <- str_replace_all(activity_week$status, "Saturday", "weekend")
activity_week$status <- str_replace_all(activity_week$status, "Sunday", "weekend")
activity_week$status <- as.factor(activity_week$status)

mean_step_imputed_ok <- activity_week %>% group_by(date, status) %>% 
        summarise(mean_imputed_yes = mean(steps, na.rm = TRUE))

plot(mean_step_imputed_ok$date[mean_step_imputed_ok$status == "weekday"], mean_step_imputed_ok$mean_imputed_yes[mean_step_imputed_ok$status == "weekday"], type = "l", col = "black", xlab = "Time", ylab = "Steps", main = "Comparing the average number of steps across days", ylim = c(0, 100))

lines(mean_step_imputed_ok$date[mean_step_imputed_ok$status == "weekend"], mean_step_imputed_ok$mean_imputed_yes[mean_step_imputed_ok$status == "weekend"], col = "blue", type= "l")

legend("topright", lty = 1, col = c("black", "blue"), legend = c("Weekday", "Weekend"), cex = 0.8, bg = "transparent")

```